Predicting trends in the quality of state-of-the-art neural networks without access to training or testing data

### 2020-09-04 submission/NatCom_all_models.tex
initial pass to modify for nat comm

%KDD% 
\vspace{-1mm}
\section{Comparing Hundreds of Models}
\label{sxn:all_cv_models}
%\vspace{-1mm}

In this section, we summarize results from a large-scale analysis of hundreds of publicly-available  models, including models developed for image 
classification, segmentation, and a range of related tasks.  
Our aim is to provide a broad analysis that complements the detailed analysis in Sections~\ref{sxn:cv} and~\ref{sxn:nlp}.
There, we considered CV and NLP models with similar structures. 
Here, we provide much broader conclusions based on a much larger set of models with a much more diverse set of structures. % about norm-based and PL-based metrics.
The models we consider have been pretrained on nine datasets.   %, shown in Table~\ref{table:datasets}.
%including ImageNet-1K, and CIFAR-10, CIFAR-100, Street View House Numbers (SVHN), Caltech-UCSD Birds-200-2011 (CUB-200-2011), Pascal VOC2012, ADE20K, Cityscapes, and Common Objects in Context (COCO). 
%%The pretrained models and their accuracy metrics are summarized in the osmr github~\cite{charles-add-link}.
We provide full details about how to reproduce these results in Appendix~\ref{sxn:appendix}.

\begin{table}[t]
\small
\begin{center}
\begin{tabular}{|p{1in}|c|c|c|c|}
%KDD% \begin{tabular}{|p{0.75in}|c|c|c|c|}
\hline
%    & Frobenius Norm & Spectral Norm & Weighted Alpha & Alpha-Norm \\
%    & $\langle\log\Vert\mathbf{W}\Vert_{F}\rangle$ & $\langle\log\Vert\mathbf{W}\Vert_{\infty}\rangle$ & $\langle\hat{\alpha}=\alpha\log\lambda_{max}\rangle$ & $\langle\log\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}\rangle$ \\
Series        & $\log\Vert\cdot\Vert_{F}$ & $\log\Vert\cdot\Vert_{\infty}$ & $\hat{\alpha}$ & $\log\Vert\cdot\Vert^{\alpha}_{\alpha}$ \\
\hline
$R^{2}$ (mean) & 0.63 &  0.55 & \textbf{0.64} & \textbf{0.64} \\
$R^{2}$ (std)  & 0.34 &  0.36 & \textbf{0.29} &          0.30 \\
\hline
$MSE$ (mean)   & 4.54 &  9.62 &          3.14 & \textbf{2.92} \\
$MSE$ (std)    & 8.69 & 23.06 &          5.14 & \textbf{5.00} \\
\hline
\end{tabular}
\end{center}
\caption{Comparison of linear regression fits for different average Log Norm and Weighted Alpha metrics across 5 CV datasets, 17 architectures, covering 108 (out of over 400) different pretrained DNNs.  
         We include regressions only for architectures with five or more data points, and which are positively correlated with test error.
         These results can be readily reproduced using the Google Colab notebooks (see Appendix~\ref{sxn:appendix}). 
        }
\label{table:results}
\end{table}

We choose ordinary least squares (OLS) regression to quantify the relationship between quality metrics (computed with the \emph{WeightWatcher} tool) 
and the reported test error and/or accuracy metrics.
We regress the metrics on the Top1 (and Top5) reported errors (as dependent variables).
These include Top5 errors for the ImageNet-1K model, percent error for the CIFAR-10/100, SVHN, CUB-200-2011 models, and Pixel accuracy
(Pix.Acc.) and Intersection-Over-Union (IOU) for other models.
We regress them individually on each of the norm-based and PL-based metrics, as described in Section~\ref{sxn:cv}.

Our results are summarized in Table~\ref{table:results}.
For the mean, larger $R^{2}$ and smaller $MSE$ are desirable; and for the standard deviation, smaller values are desirable.
Taken as a whole, over the entire corpus of data, PL-based metrics are somewhat better for both the $R^{2}$ mean and standard deviation;
and PL-based metrics are much better for $MSE$ mean and standard deviation.
These (and other) results suggest our conclusions from Sections~\ref{sxn:cv} and~\ref{sxn:nlp} hold much more generally, and they suggest
obvious questions for future work.

>>>>

\subsection{Comparing Hundreds of Models}
\label{sxn:all_cv_models}

We summarize results from a large-scale analysis of hundreds of publicly-available  models.
This broader analysis on a much larger set of CV and NLP models, with a more diverse set of architectures, and developed for a wider range of tasks, complements the previous more detailed analysis on CV and NLP models.
See Appendix~\ref{sxn:appendix} for details.
To quantify the relationship between quality metrics and the reported test error and/or accuracy metrics, we use ordinary least squares to regress the metrics on the Top1 (and Top5) reported errors (as dependent variables).
These include Top5 errors for the ImageNet-1K model, percent error for the CIFAR-10/100, SVHN, CUB-200-2011 models, and Pixel accuracy (Pix.Acc.) and Intersection-Over-Union (IOU) for other models.
We regress them individually on each of the norm-based and PL-based metrics.

\begin{table}[t]
\small
\begin{center}
\begin{tabular}{|p{1in}|c|c|c|c|}
%KDD% \begin{tabular}{|p{0.75in}|c|c|c|c|}
\hline
%    & Frobenius Norm & Spectral Norm & Weighted Alpha & Alpha-Norm \\
%    & $\langle\log\Vert\mathbf{W}\Vert_{F}\rangle$ & $\langle\log\Vert\mathbf{W}\Vert_{\infty}\rangle$ & $\langle\hat{\alpha}=\alpha\log\lambda_{max}\rangle$ & $\langle\log\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}\rangle$ \\
Series        & $\log\Vert\cdot\Vert_{F}$ & $\log\Vert\cdot\Vert_{\infty}$ & $\hat{\alpha}$ & $\log\Vert\cdot\Vert^{\alpha}_{\alpha}$ \\
\hline
$R^{2}$ (mean) & 0.63 &  0.55 & \textbf{0.64} & \textbf{0.64} \\
$R^{2}$ (std)  & 0.34 &  0.36 & \textbf{0.29} &          0.30 \\
\hline
$MSE$ (mean)   & 4.54 &  9.62 &          3.14 & \textbf{2.92} \\
$MSE$ (std)    & 8.69 & 23.06 &          5.14 & \textbf{5.00} \\
\hline
\end{tabular}
\end{center}
\caption{Comparison of linear regression fits for different average Log Norm and Weighted Alpha metrics across 5 CV datasets, 17 architectures, covering 108 (out of over 400) different pretrained DNNs.  
         We include regressions only for architectures with five or more data points, and which are positively correlated with test error.
         These results can be readily reproduced using the Google Colab notebooks. (See Appendix~\ref{sxn:appendix} for details.) 
        }
\label{table:results}
\end{table}

Results are summarized in Table~\ref{table:results}.
For the mean, larger $R^{2}$ and smaller $MSE$ are desirable; and for the standard deviation, smaller values are desirable.
Taken as a whole, over the entire corpus of data, PL-based metrics are somewhat better for both the $R^{2}$ mean and standard deviation;
and PL-based metrics are much better for $MSE$ mean and standard deviation.
These and other results suggest our conclusions hold much more generally.

### 2020-09-04 submission/NatCom_appendix.tex
initial pass to modify for nat comm

>>>>

\subsection{Additional Details on Layer Weight Matrices}

Recall that we can express the objective/optimization function for a typical DNN with $L$ layers, with activation functions $h_{l}(\cdot)$, and with $N\times M$ weight matrices $\mathbf{W}_{l}$ and biases $\mathbf{b}_{l}$, as:
\begin{equation*}
E_{DNN} \hspace{-1mm} = \hspace{-1mm} h_{L}(\mathbf{W}_{L}\cdot h_{L-1}(\mathbf{W}_{L-1}\cdot h_{L-2}(\cdots)+\mathbf{b}_{L-1})+\mathbf{b}_{L})  .
\end{equation*}
We expect that most well-trained, production-quality models will employ one or more forms of regularization, such as Batch Normalization (BN), Dropout, etc., and many will also contain additional structure such as Skip Connections, etc. 
Here, we will ignore these details, and will focus only on the pretrained layer weight matrices $\mathbf{W}_{l}$.
Typically, this model would be trained on some labeled data $\{d_{i},y_{i}\}\in\mathcal{D}$, using Backprop, by minimizing the loss $\mathcal{L}$.
For simplicity, we do not indicate the structural details of the layers (e.g., Dense or not, Convolutions or not, Residual/Skip Connections, etc.). 
Each layer is defined by one or more layer 2D weight matrices $\mathbf{W}_{l}$, and/or the 2D feature maps $\mathbf{W}_{l,i}$ extracted from 2D Convolutional (Conv2D) layers.
A typical modern DNN may have anywhere between 5 and 5000 2D layer matrices.

For each Linear Layer, we get a  single $(N\times M)$ (real-valued) 2D weight matrix, denoted $\mathbf{W}_{l}$, for layer $l$.  
This includes Dense or Fully-Connected (FC) layers, as well as 1D Convolutional (Conv1D) layers, Attention matrices, etc.
We ignore the bias terms $\mathbf{b}_{l}$ in this analysis. 
Let the aspect ratio be $Q=\frac{N}{M}$, with $Q\ge 1$.
For the Conv2D layers, we have a 4-index Tensor, of the form $(N\times M \times c\times d)$, consisting
of $c\times d$ 2D feature maps of shape $(N\times M)$.    
We  extract $n_{l}=c\times d$ 2D weight matrices $\mathbf{W}_{l,i}$, one for each feature map $i=[1,\dots,n_{l}]$ for layer $l$.

\subsection{Additional Details on HT-SR Theory}

The original work on HT-SR Theory~\cite{MM18_TR,MM19_HTSR_ICML,MM20_SDM} considered NNs including AlexNet and InceptionV3 (as well as DenseNet, ResNet, and VGG), and it showed that for nearly every $\mathbf{W}$, the (bulk and tail) of the ESDs can be fit to a truncated PL and the PL exponents $\alpha$ nearly all lie within the range $\alpha\in(1.5,5)$.
Our meta-analysis, the main results of which are summarized in this paper, has shown that these results are ubiquitous.
For example, 
upon examining nearly 10,000 layer weight matrices $\mathbf{W}_{l,i}$ across hundreds of different modern pre-trained DNN architectures, the ESD of nearly every $\mathbf{W}$ layer matrix can be fit to a truncated PL:
$70-80\%$ of the time, the fitted PL exponent $\alpha$ lies in the range $\alpha\in(2,4)$; and  
$10-20\%$ of the time, the fitted PL exponent $\alpha$ lies in the range $\alpha< 2$.  
Of course, there are exceptions: in any real DNN, the fitted $\alpha$ may range anywhere from $\sim 1.5$ to $10$ or higher (and, of course, larger values of $\alpha$ may indicate that the PL is not a good model for the data).  
Still, overall, in nearly all large, pre-trained DNNs, the correlations in the  weight matrices exhibit a remarkable Universality, being both Heavy Tailed, and having small---but not too small---PL exponents.
